---
title: "SpatialData.Rmd"
author: "Witold Wolski"
date: "May 5, 2017"
output: html_document
---

```{r}
install.packages(c("sp","plotGoogleMaps","spatstat","raster","maptools","plotrix","rgeos"))
```

```{r setup, include=FALSE}
library(sp)
library(plotGoogleMaps)  
library(spatstat)
library(raster)
library(maptools)
library(plotrix)
library(rgeos)

```

```{r eval=FALSE}
data <- read.csv("http://www.fabioveronesi.net/Blog/2014-05-metropolitan-street.csv")  
head(data)
write.csv(data, file="Data/OriginalDataOfFabio.csv")
```

```{r}
data <- read.csv("Data/OriginalDataOfFabio.csv")
str(data)
data <- data[!is.na(data$Longitude)&!is.na(data$Latitude),]  
```

#Point pattern analysis

A point process is a stochastic process for which we observe its results, or events, only in a specific region, which is the area under study, or simply window. The location of the events is a point pattern (Bivand et al., 2008).
In R the package for Point Pattern Analysis is spatstat, which works with its own format (i.e. ppp). There are ways to transform a data.frame into a ppp object, however in this case we have a problem. The crime dataset contains lots of duplicated locations. We can check this by first transform data into a SpatialObject and then use the function zerodist to check for duplicated locations:

```{r}
coordinates(data)=~Longitude+Latitude  
zero <- zerodist(data)
class(zero)
length(unique(zero[,1]))  
```

#Loading Natural Earth Provinces dataset to define window for Point Pattern Analysis  

There is also another potential issue with this dataset. Even though the large majority of crimes are reported for London, some of them (n=660) are also located in other areas. Since these crimes are a small fraction of the total I do not think it makes much sense to include them in the analysis, so we need to remove them. To do so we need to import a shapefile with the borders of the Greater London region. Natural Earth provides this sort of data, since it distributes shapefiles at various resolution. For this analysis we would need the following dataset: Admin 1 – States, Provinces

To download it and import it in R we can use the following lines:


```{r eval=FALSE}
download.file("http://www.naturalearthdata.com/http//www.naturalearthdata.com/download/10m/cultural/ne_10m_admin_1_states_provinces.zip",destfile="ne_10m_admin_1_states_provinces.zip")  
unzip("ne_10m_admin_1_states_provinces.zip",exdir="NaturalEarth") 
```

```{r}
border <- shapefile("NaturalEarth/ne_10m_admin_1_states_provinces.shp")  
```
    
To extract only the border of the Greater London regions we can simply subset the SpatialPolygons object as follows:

```{r}
GreaterLondon <- border[paste(border$region)=="Greater London",]  
```

Now we need to overlay it with crime data and then eliminate all the points that do not belong to the Greater London region. To do that we can use the following code:

```{r}
#Extract crimes in London  
projection(data)=projection(border)  
overlay <- over(data,GreaterLondon)  

data$over <- overlay$OBJECTID_1  

data.London <- data[!is.na(data$over),]  
```

#Simple Plot

```{r}
#jpeg("PP_plot.jpg",2500,2000,res=300)  
plot(data.London,pch="+",cex=0.5,main="",col=data.London$Crime.type)  
plot(GreaterLondon,add=T)  
legend(x=-0.53,y=51.41,pch="+",col=unique(data.London$Crime.type),legend=unique(data.London$Crime.type),cex=0.4)  
#dev.off()  
```


#Summary statistics for point patterns
#The coordinates of the mean center are simply the mean value of X and Y
#therefore we can use the function mean() to determine their value

```{r}
mean_centerX <- mean(data.London@coords[,1])
mean_centerY <- mean(data.London@coords[,2])

```

#Similarly we can use the function sd() to determine the standard deviation of X and Y

```{r}
standard_deviationX <- sd(data.London@coords[,1])
standard_deviationY <- sd(data.London@coords[,2])
```

#This is the formula to compute the standard distance

```{r}
standard_distance <- sqrt(sum(((data.London@coords[,1]-mean_centerX)^2+(data.London@coords[,2]-mean_centerY)^2))/(nrow(data.London)))  
```


```{r}
plot(data.London,pch="+",cex=0.5,main="")  
plot(GreaterLondon,add=T)  
points(mean_centerX,mean_centerY,col="red",pch=16)  
draw.circle(mean_centerX,mean_centerY,radius=standard_distance,border="red",lwd=2)  

```



```{r}
plot(data.London,pch="+",cex=0.5,main="")  
plot(GreaterLondon,add=T)  
points(mean_centerX,mean_centerY,col="red",pch=16)  
draw.ellipse(mean_centerX,mean_centerY,a=standard_deviationX,b=standard_deviationY,border="red",lwd=2)  

```




#Working with spatstat  
```{r}
Drugs <- data.London[data.London$Crime.type==unique(data.London$Crime.type)[3],]  
Drugs <- remove.duplicates(Drugs)  

```


#Transform GreaterLondon in UTM  

```{r}
GreaterLondonUTM <- spTransform(GreaterLondon,CRS("+init=epsg:32630"))  
Drugs.UTM <- spTransform(Drugs,CRS("+init=epsg:32630"))  

```

#Transforming the SpatialPolygons object into an owin object for spatstat, using a function in maptools  
```{r}
window <- as.owin(GreaterLondonUTM)  
```


#Now we can extract one crime and   
```{r}
Drugs.ppp <- ppp(x=Drugs.UTM@coords[,1],y=Drugs.UTM@coords[,2],window=window)  
```

#Calculate Intensity  
```{r}
Drugs.ppp$n/sum(sapply(slot(GreaterLondonUTM, "polygons"), slot, "area"))  
```

#Alternative approach  
```{r}
summary(Drugs.ppp)$intensity  
```


#Quadrat counting Intensity  
```{r}
plot(Drugs.ppp,pch="+",cex=0.5,main="Drugs")  
plot(quadratcount(Drugs.ppp, nx = 4, ny = 4),add=T,col="red")  
```

#Intensity by Borough
```{r}
Local.Intensity <- data.frame(Borough=factor(),Number=numeric())  
for(i in unique(GreaterLondonUTM$name)){  
sub.pol <- GreaterLondonUTM[GreaterLondonUTM$name==i,]  

sub.ppp <- ppp(x=Drugs.ppp$x,y=Drugs.ppp$y,window=as.owin(sub.pol))  
Local.Intensity <- rbind(Local.Intensity,data.frame(Borough=factor(i,levels=GreaterLondonUTM$name),Number=sub.ppp$n))  
}  
```



```{r}
colorScale <- color.scale(Local.Intensity[order(Local.Intensity[,2]),2],color.spec="rgb",extremes=c("green","red"),alpha=0.8)  

par(mar=c(5,13,4,2))   
barplot(Local.Intensity[order(Local.Intensity[,2]),2],names.arg=Local.Intensity[order(Local.Intensity[,2]),1],horiz=T,las=2,space=1,col=colorScale)  

```


#Kernel Density (from: Baddeley, A. 2008. Analysing spatial point patterns in R)   
#Optimal values of bandwidth  
```{r}
bw.diggle(Drugs.ppp)  
bw.ppl(Drugs.ppp)  
bw.scott(Drugs.ppp)  
```

#Plotting  
```{r}
par(mfrow=c(2,2))  
plot(density.ppp(Drugs.ppp, sigma = bw.diggle(Drugs.ppp),edge=T),main=paste("h =",round(bw.diggle(Drugs.ppp),2)))  
plot(density.ppp(Drugs.ppp, sigma = bw.ppl(Drugs.ppp),edge=T),main=paste("h =",round(bw.ppl(Drugs.ppp),2)))  
plot(density.ppp(Drugs.ppp, sigma = bw.scott(Drugs.ppp)[2],edge=T),main=paste("h =",round(bw.scott(Drugs.ppp)[2],2)))  
plot(density.ppp(Drugs.ppp, sigma = bw.scott(Drugs.ppp)[1],edge=T),main=paste("h =",round(bw.scott(Drugs.ppp)[1],2)))  
```


#G Function  
```{r}
plot(Gest(Drugs.ppp),main="Drug Related Crimes")  
```

